
---
title: "Determining System Information"
author: "A. Sanchez Meador"
date: "`r Sys.Date()`"
output: html_document
---
---

## Introduction

This document provides a way to gather and display information about your system's hardware. Specifically, it determines:

- The number of logical and physical processors
- The total amount of RAM and available RAM
- The speed of the processor

## Prerequisites

To run this, you'll need the `parallel` library.

```{r}
library(parallel)
```

## Determine Number of Processors

Here, we determine both the logical (which includes hyper-threading if available) and physical cores of the processor.

```{r processor-info}
logical_cores <- detectCores(logical = TRUE)
physical_cores <- detectCores(logical = FALSE)

cat("Number of logical cores:", logical_cores, "
")
cat("Number of physical cores:", physical_cores, "
")
```

## Determine RAM Information

Depending on the operating system (Windows vs. Unix-based), the method to determine RAM differs. For Windows, we use the `systeminfo` command, while for Unix-based systems, we read from `/proc/meminfo`.

```{r ram-info}
if (.Platform$OS.type == "windows") {
  
  sysinfo <- system("systeminfo", intern = TRUE)
  
  # Extract total RAM from systeminfo output
  total_ram_line <- grep("Total Physical Memory:", sysinfo, value = TRUE)
  num_string_total <- gsub(".*?([0-9,]+) MB.*", "\\1", total_ram_line)
  total_ram <- as.numeric(gsub(",", "", num_string_total)) / 1024  # Convert to GB
  
  # Extract available RAM from systeminfo output
  free_ram_line <- grep("Available Physical Memory:", sysinfo, value = TRUE)
  num_string_free <- gsub(".*?([0-9,]+) MB.*", "\\1", free_ram_line)
  free_ram <- as.numeric(gsub(",", "", num_string_free)) / 1024  # Convert to GB
  
} else { # Unix-based systems (e.g., Linux, macOS)
  
  meminfo <- readLines("/proc/meminfo")
  total_ram_line <- grep("MemTotal:", meminfo, value = TRUE)
  free_ram_line <- grep("MemFree:", meminfo, value = TRUE)
  
  # Extract the values from the lines
  total_ram <- as.numeric(gsub(".*?([0-9]+) kB", "\\1", total_ram_line)) / 1024^2  # Convert to GB
  free_ram <- as.numeric(gsub(".*?([0-9]+) kB", "\\1", free_ram_line)) / 1024^2  # Convert to GB
  
}

cat("Total RAM (GB):", total_ram, "\n")
cat("Available RAM (GB):", free_ram, "\n")
```

## Determine Processor Speed

Like the RAM information, the method to determine the processor speed is dependent on the operating system.

```{r cpu-speed}
if (.Platform$OS.type == "windows") {
  cpu_speed <- system("wmic cpu get MaxClockSpeed", intern=TRUE)[2]
  cat("Processor speed (MHz):", cpu_speed, "
")
} else { # Unix-based systems
  cpu_speed <- system("sysctl -n hw.cpufrequency", intern=TRUE)
  cpu_speed <- as.numeric(cpu_speed) / 10^6 # Convert to MHz
  cat("Processor speed (MHz):", cpu_speed, "
")
}
```

## Conclusion

This document offers a concise way to gather key hardware details about your system. You can easily extend this to include more details as required.
